---
title: "Data607 Web APIs"
author: "Michael Muller"
date: "April 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(RCurl)
library(jsonlite)
library(knitr)
```

### Step 1 : Apply for a key (Many APIs will ask for authenticators, but this restful API just asks for a user specific key inside your http request.)
### Step 2 : Select from the NY times API tree, which category you would like to request information for. I chose movie reviews!
### Step 3 : I use a combination of RCurl and jsonlite to retrieve and "jsonify" the object I recieve from NY times.
### Step 4 : I seperate all three tables including in my object, into three different dataframes.
#### Bonus Step : If I were to expand on this dataframe, I would use scrapeR package to retrieve the links from the movieLinks dataframe, then use xpathSAPPLY() to retrieve genres from the review links, and add that vector to movies dataframe.

```{r}
apiKey = "?api-key=5663a997971b47b1946e2c156218c2b8"
baseUrl = "https://api.nytimes.com/svc/movies/v2/reviews/search.json"
myUrl = paste(baseUrl,apiKey,sep="")
data = getURL(myUrl)
json =fromJSON(data)
# Make dataframe
movies = data.frame(json$results)
# Seperate dataframes
movieLinks = movies$link
movieMultimedia = movies$multimedia
movies$multimedia = NULL
movies$link = NULL
# Rename columns
names(movies) = c('Title','Rating','Suggested','Critic','Headline','Summary','Publication date','Release date','Updated date')
# Change binary results to plain english
movies$Suggested[movies$Suggested==1] = 'Yes'
movies$Suggested[movies$Suggested==0] = 'No'

```

```{r results='asis',error=F,warning=F}

kable(head(movies),format="markdown")
```

